Capturing subdiffusive solute dynamics and predicting selectivity in nanoscale pores with time series modeling

### 2020-01-15 Ben_Manuscripts/stochastic_transport/Final_Draft.tex
Updated tex with MRS7 comments.
Simplifying solute flux section

%BJC5: wondering about replacing "Predicts long timescale behavior" with "predicts selectivity"
%MRS6: I would focus first on the outputs.  Not sure about the behavior/vs. selectivity.  Probably 
% selelctivity since it's more results oriented. ``of varying complexity'' is unneeded in the titel
%\title{Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics
%and Predicts Long Timescale Behavior in Nanoscale Pores.}
% BJC6: The only issue with selectivity is that I don't think it would include a timescale.
\title{Capturing subdiffusive solute dynamics and predicting long timescale selectivity in nanoscale pores
with time series modeling}
\author{Benjamin J. Coscia \and Michael R. Shirts}

@@

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore 2 variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies 2 sets of parameters based on the solute's radial distance from the
  closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which has separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that our models reasonably reproduce
  the MSDs calculated from MD simulations. Finally, we demonstrate how one can
  use our models to estimate flux of a solute across a macroscopic-length pore
  and, based on those quantities, the membrane's selectivity towards each solute.
  Overall, this work helps to connect microscopic chemically-dependent solute 
  motions and macroscopic membrane performance.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that can be 
  cross-linked into mechanically strong and highly selective nanoporous membranes.~\cite{gin_polymerized_2008}
  They may provide a promising alternative to conventional membrane separation techniques
  by being selective based not only on solute size and charge, but on solute chemical 
  functionality as well.~\cite{dischinger_application_2017} One can tune the functionality
  of LLC monomers in order to enhance or weaken specific solute-membrane interactions.~\cite{dischinger_effect_2017}
  Two general types of LLC membranes are being actively developed. The inverted hexagonal, 
  H\textsubscript{II}, LLC phase has densely packed, uniform-sized pores on the order of 1 nm
  in size. A perfectly synthesized H\textsubscript{II} phase has an ideal geometry for high
  throughput separations, however aligning the microscopic hexagonal mesophases on a large 
  scale has shown limited success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLCs share the uniform and nm-sized chemically complex pores of
  the H\textsubscript{II} phase but its geometry consists of a tortuous network of 3D 
  interconnected pores which do not require alignment.~\cite{carter_glycerol-based_2012} While 
  its pore structure may decrease a Q\textsubscript{I} membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to synthesize at scale.

@@

  In our previous work, we used molecular dynamics (MD) simulations to study the transport
  of 20 small polar molecules in an H\textsubscript{II} phase LLC membrane.~\cite{coscia_chemically_2019}
  %MRS6: sentence seems out of place here? True, but does it support the thesis here?
  %BJC6: I said above how the QI phase is easier to synthesize, so I felt it necessary to 
  % quickly justify why we studied the HII phase. Reworded to make that clearer. I'm also open to
  % eliminating the sentence.
  We chose to study the H\textsubscript{II} phase because it is a simpler geometry to model than 
  the Q\textsubscript{I} phase.  
  In general, we observed subdiffusive transport behavior characterized by intermittent hops 
  separated by periods of entrapment. We identified three mechanisms responsible for the 
  solute trapping behavior: entanglement among monomer tails, hydrogen bonding with monomer
  head groups, and association with the monomer's sodium counter ions.

  %BJC4: Added next two paragraphs to help ease more into time series modeling
  So far, our molecular models have provided valuable qualitative mechanistic insight 
  with some quantitative support which already allows us to speculate about new LLC
  monomer designs. However, they would be of greater value to a larger set of 
  researchers if we could provide quantitative predictions of macroscopic observables
  such as solute flux and selectivity. Due to the size of the types of systems we are 
  studying (at least 62,000 atoms) it is prohibitively expensive and time-consuming to
  run simulations longer than those performed for this work ($\sim$5 $\mu$s). 
  Even with the relatively large size of our system, we only simulate 24 solute trajectories per unit cell in order
  to minimize solute-solute interactions. This results in relatively high uncertainties
  in observables such as mean squared displacement (MSD), preventing reliable long
  timescale predictions. We are in need of a way of studying the collective motion
  of a much larger set of solutes over much longer timescales.

  %BJC4: 1st paragraph moved to here
  Mathematical descriptions of transport in complex separations membranes are 
  a powerful way to understand mechanisms and formulate design principles. 
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  The complexity of a well-fit model generally parallels the complexity of the transport 
  mechanism at hand, as well as the transport information the model conveys.
  In dense homogeneous membranes, the solution-diffusion model can extract 
  diffusion and partition coefficients and has successfully predicted solute 
  transport rates.~\cite{wijmans_solution-diffusion_1995} Analogously, pore-flow
  models yield predictions of diffusion coefficients and solute transport rates 
  in nanoporous membranes.~\cite{paul_diffusive_1974} Modern single particle 
  tracking approaches have taken researchers beyond continuum modeling allowing them to 
  characterize complex diffusive behavior.~\cite{manzo_review_2015} At the molecular
  level, one can use molecular dynamics (MD) simulations to study both single 
  particle dynamics and bulk transport properties with atomic-level insight.~\cite{coscia_chemically_2019,maginn_best_2018}
  All of these approaches facilitate generation of hypotheses about the
  molecular origins of separations by attempting to give a more intuitive 
  understanding of how solutes move as a function of their environment, in turn
  suggesting experiments that could be performed.

  %BJC4: Tried to make this paragraph connect to previous a little better
  Using a bottom-up modeling approach, we can parameterize single particle behavior
  by extracting their trajectories from MD simulations and studying the properties that
  have the greatest influence on solute dynamics. With this information, we can learn
  how to construct an ensemble of characteristic single solute trajectories with these
  properties which would be useful for making long timescale predictions with computational
  ease and lower uncertainties. Fortunately, there is an abundance of information encoded
  in the complex solute trajectories. Into a
  model, we can incorporate fluctuations and time-dependent correlations in solute
  position as well as the length of time solutes are trapped. We can add further
  detail by integrating our time series analysis with our knowledge of the primary
  trapping mechanisms as well as the solute's changing chemical environment within
  the heterogeneous membrane.

@@

  %MRS4: consider more citations below, when you talk about what researchers have done.
  %BJC4: added citations of prominent anomalous diffusion work
  Researchers have developed a rigorous theoretical foundation for describing the 
  motion of particles exhibiting non-Brownian, or anomalous, transport behavior.
  \cite{metzler_random_2000,bouchaud_anomalous_1990} The tools introduced by 
  fractional calculus are instrumental to this theory.~\cite{gorenflo_fractional_1997}
  They allow us to generalize the normally linear diffusion equation to fractional
  derivative orders, providing descriptions of a much more diverse set of behavior.~\cite{klages_anomalous_2008}

@@

  %BJC4: added mention of the anomalous subdiffusion model we chose 
  Our first modeling approach is based on the anomalous diffusion literature, 
  applying the mathematical formalism describing subdiffusive transport to observed
  solute trajectories. Specifically, we treated our system as an FBM process subordinate
  to a CTRW, or subordinated FBM (sFBM) for short. We fit sFBM
  models in two ways: First, we parameterized solute motion using a single set of parameters fit
  to a hypothesized anomalous diffusion model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's radial
  distance from the closest pore center. This allows us to include different dynamical
  behavior of solutes observed within the pore region versus within the monomer tails.

  Where applicable, the sFBM model is an elegant way to characterize single 
  particle trajectories and in addition to providing an easy and efficient way to 
  simulate solute trajectories, can help uncover the underlying solute-membrane
  interactions which result in observed transport mechanisms. It can help 
  researchers formulate hypotheses to explain why a certain set of parameters is
  characteristic to a specific solute. Once one has knowledge of the primary 
  mechanisms leading to anomalous diffusion behavior, one may gain additional
  insight by formulating a state-based model.

  % moved parts of method to here.
  Markov state models (MSMs) are a popular class of models used to project long timescale
  system properties based on molecular simulation trajectories by identifying
  different dynamical modes and quantifying the rates of transitions between them.
  MSMs are frequently used to study systems with slow dynamics, such as protein 
  folding.~\cite{snow_how_2005,chodera_automatic_2007} Researchers typically aim to 
  come up with a low dimensional representation of the system based on features 
  which preserve the process kinetics. This still often results in hundreds to thousands
  of distinct states.~\cite{chodera_markov_2014}

  Our second modeling approach applies an extended MSM framework to a relatively 
  small set of known states based on the three previously observed solute trapping mechanisms.  
  Standard MSMs are typically applied to determine equilibrium populations
  of and the kinetics of transitions between various state configurations.~\cite{bowman_using_2009} 
  We extend the framework to include state-dependent fluctuations and correlations in 
  solute position. The magnitude of solute fluctuations from its average trapped 
  position and the degree of correlation with previous fluctuations is 
  determined by the current state. To distinguish our approach from standard MSMs, we have
  named it the Markov state-dependent dynamical model (MSDDM).

  %BJC4: making goals more explicit
  We determine the degree of success of our modeling approaches in two ways. First,
  if we can closely reproduce solute MSDs measured from MD simulations with realizations
  of our models, then it is likely that the model sufficiently captures solute 
  dynamics and can be used in a predictive capacity. Even if a model fails to 
  reproduce the MD MSDs, there is value in uncovering the cause of the deviation.
  The second measure of success is by analysis of the parameters defining each 
  model. The parameters themselves should provide clues that help explain solute 
  behavior in a coherent way. A model that successfully reproduces MD MSDs, but 
  with nonsensical parameters has severely diminished value as it could be a 
  consequence of luck.

@@

  We evaluate the two modeling approaches by using them to characterize the dynamical
  behavior of the four fastest moving solutes studied in our previous work.
  Specifically, we study methanol, urea, ethylene glycol and acetic acid. In addition
  to moving quickly, allowing them to extensively explore membrane structural space,
  these solutes have a range of chemical functionality and experience each of the 
  three trapping mechanisms to different extents.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an atomistic
  molecular model of four pores in a monoclinic unit cell with 10\% water by weight 
  (see Figure~\ref{fig:membrane_structure}). Approximately one third of the water 
  molecules occupy the tail region with the rest near the pore center. We chose to
  focus our study on the 10 wt\% water system because solutes move significantly 
  faster than in the 5 wt\% system studied previously. Appropriate stochastic 
  modeling requires that solutes sample the accessible mechanisms with representative
  probability.

@@

  We chose to study a subset of the 4 fastest moving solutes from our previous
  work: methanol, acetic acid, urea and ethylene glycol. For each solute we 
  created a separate system and to each system we added 6 solutes per pore 
  for a total of 24 solutes. This number of solutes per pore provides a balance
  of a low degree of interaction between solutes, since we are primarily interested 
  in solute-membrane interactions at present, as well as a sufficient amount of 
  data from which to generate statistics on the time scales which we simulate.
  Further details on the setup and equilibration of these systems are described in
  our previous work.\cite{coscia_chemically_2019}

  We extended the 1 $\mu$s simulations of our previous work to 5 $\mu$s in order
  to collect ample data. We simulated the system with a time step of 2 fs at a pressure
  of 1 bar and 300 K controlled by the Parrinello-Rahman barostat and the v-rescale 
  thermostat respectively. We recorded frames every 0.5 ns.

@@

  %MRS6: next paragraph could be clearer.  Maybe motivate at the beginning why the importance of time. vs. ensemble averated. It's sprung a sort of a surprise there's a difference 
  %BJC6: reorganized a little and broke into paragraphs meant to convey separate ideas.

  One can measure MSD in two ways. The time-averaged MSD measures all observed
  displacements over time lag $\tau$: 
  \begin{equation}
  \overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. The ensemble averaged MSD measures
  displacements with respect to a particle's initial position:
  \begin{equation}
  \langle z^2(t) \rangle = \langle z(t) - z(0) \rangle^2
  \label{eqn:ensemble_msd}
  \end{equation}
  Fits to the \textit{ensemble} averaged MSD will always reproduce the form of 
  Equation~\ref{eqn:msd_form}.

@@

  \textit{Dwell time distributions:} For subdiffusive transport, the distribution 
  of dwell times is expected to fit a power law distribution 
  proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  Because we are limited to taking measurements at discrete intervals dictated by the output 
  frequency of our simulation trajectories, we fit the empirical dwell times
  to a discrete power law distribution whose maximum likelihood $\alpha$ 
  parameter we calculated by maximizing the likelihood function: 
  %BJC6: I could move these MLE equations to Supporting
  \begin{equation}
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, t_{min}) -
	\beta\sum_{i=1}^{n} \ln t_i 
  \label{eqn:powerlaw_likelihood}
  \end{equation}
  where $\beta = 1 + \alpha$, $t_i$ are collected dwell time data points,
  $n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
  where $t_{min}$ is the smallest measured value of	$t$.~\cite{clauset_power-law_2009}

@@

  \textit{Correlated hop length distributions:} The marginal distribution of hops by solutes 
  undergoing an sFBM process is Gaussian, therefore we parameterize it by its standard deviation, 
  $\sigma$.~\cite{metzler_random_2000, metzler_anomalous_2014,neusius_subdiffusion_2009}
  The measured mean is always very close to zero so we assume that it is exactly zero
  in our time series simulations since we have no reason to expect drift in either direction
  based on pore symmetry.

  % BJC3: Making a logical argument regarding how we estimate H here is important
  % because I haven't seen anyone do it this way.
  sFBM implies that hops are correlated which we describe using the Hurst parameter, $H$. 
  The autocovariance function of hop lengths has the analytical form:~\cite{mandelbrot_fractional_1968}
  \begin{equation}
    \gamma(k) = \dfrac{\sigma^2}{2}\bigg[|k-1|^{2H} - 2|k|^{2H} + |k+1|^{2H}\bigg]
  \label{eqn:fbm_autocorrelation}
  \end{equation}
  where $\sigma^2$ is the variance of the underlying Gaussian distribution from which hops are
  drawn and $k$ is the time lag, or number of increments between hops. The hop autocorrelation
  function is simply Equation~\ref{eqn:fbm_autocorrelation} normalized by the variance. 
  When $H < 0.5$, hops are negatively correlated, when $H = 0.5$ we recover Brownian motion
  and when $H > 0.5$, one observes positive correlation between hops. 
  %BJC6: I could move the following to supporting and keep last sentence.
  There are many methods
  for estimating the Hurst parameter for a time series.~\cite{clegg_practical_2006} It can
  be a difficult task because Equation~\ref{eqn:fbm_autocorrelation} decays slowly to zero, 
  especially when $H > 0.5$, meaning one needs to study large time lags with high frequency.
  Fortunately (from a mathematical perspective), all of our solutes show anti-correlated motion, so most of the information in
  Equation~\ref{eqn:fbm_autocorrelation} is contained within the first few time lags (see 
  Figure~\ref{S-fig:hurst_autocorrelation} of the Supporting Information). Any hope of
  fitting longer time lags to our data is lost in the noise since the analytical values
  are so close to zero. Therefore, we obtained H by performing a least squares fit of 
  Equation~\ref{eqn:fbm_autocorrelation} to the first 
  ten  
  0.5 ns time lags of the 
  empirically measured autocorrelation function.

@@

%    $$ A_i = \frac{J}{(\Delta P - \Delta \pi)} $$
%  $$ S = \frac{A_i}{A_j} for a given pore length, L$$
%  $$J = A(\Delta P - \Delta \pi)$$
%  $$A = \frac{P_i}{L} $$
%  $$P_i = \frac{flux}{drivingforce \cdot L} [=] \frac{particles}{s \cdot bar \cdot L}$$

  %BJC4: we have better ways of getting peaks now
%  In the absence of an analytical equation for the MFPT, we instead report the location of
%  the peak of the passage time distribution. This has the added advantage of treating
%  cases where the passage time distribution diverges and no mean exists. While we will
%  call the peak in the distribution the MFPT, it is important to recognize that this 
%  value is more representative of a typical passage time rather than a mean.

  % BJC: concentration profile stuff
%  One would then segment the trajectory
%  each time it crosses $0$ and reflect all of the negative trajectories. The ensemble of
%  trajectories is consistent with the fraction of partic

>>>>

%BJC5: wondering about replacing "Predicts long timescale behavior" with "predicts selectivity"
%MRS6: I would focus first on the outputs.  Not sure about the behavior/vs. selectivity.  Probably 
% selelctivity since it's more results oriented. ``of varying complexity'' is unneeded in the tite
%\title{Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics
%and Predicts Long Timescale Behavior in Nanoscale Pores.}
% BJC6: The only issue with selectivity is that I don't think it would include a timescale.
%MRS7: true, could just be ``predicting selectivity''
\title{Capturing subdiffusive solute dynamics and predicting selectivity in nanoscale pores
with time series modeling}
\author{Benjamin J. Coscia \and Michael R. Shirts}

@@

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore 2 variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies 2 sets of parameters based on the solute's radial distance from the
  closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which has separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. 
  %BJC7: added  
  However, qualitative differences between MD and simulated MSDDM trajectories
  may limit its usefulness. 
  Finally, we demonstrate how one can use our models to estimate flux of a solute
  across a macroscopic-length pore and, based on those quantities, the membrane's selectivity towards each solute.
  This work helps to connect microscopic chemically-dependent solute 
  motions and macroscopic membrane performance.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that can be 
  cross-linked into mechanically strong and highly selective nanoporous membranes.~\cite{gin_polymerized_2008}
  They may provide a promising alternative to conventional membrane separation techniques
  by being selective based not only on solute size and charge, but on solute chemical 
  functionality as well.~\cite{dischinger_application_2017} One can tune the functionality
  of LLC monomers in order to enhance or weaken specific solute-membrane interactions.~\cite{dischinger_effect_2017}
  Two general morphologies of LLC membranes are being actively developed. The inverted hexagonal, 
  H\textsubscript{II}, LLC phase has densely packed, uniform-sized pores on the order of 1 nm
  in size. A perfectly synthesized H\textsubscript{II} phase has an ideal geometry for high
  throughput separations, however aligning the microscopic hexagonal mesophases on a large 
  scale has shown limited success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLCs share the uniform and nm-sized chemically complex pores of
  the H\textsubscript{II} phase but its geometry consists of a tortuous network of 3D 
  interconnected pores which do not require alignment.~\cite{carter_glycerol-based_2012} While 
  its pore structure may decrease a Q\textsubscript{I} membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to synthesize at scale.~\cite{hatakeyama_nanoporous_2010}

@@

  In our previous work,~\cite{coscia_chemically_2019} we used molecular dynamics (MD) 
  simulations to study the transport of 20 small polar molecules in an H\textsubscript{II}
  phase LLC membrane.
  %MRS6: sentence seems out of place here? True, but does it support the thesis here?
  %BJC6: I said above how the QI phase is easier to synthesize, so I felt it necessary to 
  % quickly justify why we studied the HII phase. Reworded to make that clearer. I'm also open to
  % eliminating the sentence.
  %MRS6: It was also becasue they thought it was more likely to be interesting initially.  I'm in favor of 
  % cutting for simplicity. 
  We chose to study the H\textsubscript{II} phase because it is a simpler geometry to model than 
  the Q\textsubscript{I} phase.  
  In general, we observed subdiffusive transport behavior characterized by intermittent hops 
  separated by periods of entrapment. We identified three mechanisms responsible for the 
  solute trapping behavior: entanglement among monomer tails, hydrogen bonding with monomer
  head groups, and association with the monomer's sodium counter ions.

  %BJC4: Added next two paragraphs to help ease more into time series modeling
  Up through our previous studies, our molecular models have provided valuable 
  qualitative mechanistic insight with some quantitative support. This insight
  already allows us to speculate about new LLC monomer designs. However, they 
  would be of greater value to a larger set of researchers if we could provide 
  quantitative predictions of macroscopic observables such as solute flux and 
  selectivity. Due to the size of the types of systems we are studying (at least
  (62,000 atoms) it is prohibitively expensive and time-consuming to
  run simulations longer than those performed for this work ($\sim$5 $\mu$s). 
  Even with the relatively large size of our system, we only simulate 24 solute trajectories per unit cell in order
  to minimize solute-solute interactions. This results in relatively high uncertainties
  in observables such as mean squared displacement (MSD), preventing reliable long
  timescale predictions. We are in need of a way of studying the collective motion
  of a much larger set of solutes over much longer timescales.

  %BJC4: 1st paragraph moved to here
  Mathematical descriptions of transport in complex separations membranes are 
  a powerful way to understand mechanisms and formulate design principles. 
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  The complexity of a well-fit model generally parallels the complexity of the transport 
  mechanism being studied, as well as the transport information the model conveys.
  In dense homogeneous membranes, the solution-diffusion model can extract 
  diffusion and partition coefficients and has successfully predicted solute 
  transport rates.~\cite{wijmans_solution-diffusion_1995} Analogously, pore-flow
  models yield predictions of diffusion coefficients and solute transport rates 
  in nanoporous membranes.~\cite{paul_diffusive_1974} Modern single particle 
  tracking approaches have taken researchers beyond continuum modeling allowing them to 
  characterize complex diffusive behavior.~\cite{manzo_review_2015} At the molecular
  level, one can use molecular dynamics (MD) simulations to study both single 
  particle dynamics and bulk transport properties with atomic-level insight.~\cite{coscia_chemically_2019,maginn_best_2018}
  All of these approaches facilitate generation of hypotheses about the
  molecular origins of separations by attempting to give a more intuitive 
  understanding of how solutes move as a function of their environment, in turn
  suggesting experiments that could be performed.

  %BJC4: Tried to make this paragraph connect to previous a little better
  Using a bottom-up modeling approach, we can parameterize single particle behavior
  by extracting their trajectories from MD simulations and studying the properties that
  have the greatest influence on solute dynamics. With this information, we can learn
  how to construct an ensemble of characteristic single solute trajectories with these
  properties which would be useful for making long timescale predictions with computational
  ease and lower uncertainties. Fortunately, there is an abundance of information encoded
  in the complex solute trajectories. We can incorporate fluctuations and time-dependent 
  correlations in solute position as well as the length of time solutes are trapped into a
  model. We can add further
  detail by integrating our time series analysis with our knowledge of the primary
  trapping mechanisms as well as the solute's changing chemical environment within
  the heterogeneous membrane.

@@

  %MRS4: consider more citations below, when you talk about what researchers have done.
  %BJC4: added citations of prominent anomalous diffusion work
  %BJC7: changed thesis to focus more on this paper
  We constructed our first model by applying the rigorous theoretical foundation 
  already developed to describe the motion of particles exhibiting non-Brownian, or
  anomalous, transport behavior.~\cite{metzler_random_2000,bouchaud_anomalous_1990}
  The tools introduced by fractional calculus are instrumental to this theory.~\cite{gorenflo_fractional_1997}
  They allow us to generalize the normally linear diffusion equation to fractional
  derivative orders, providing descriptions of a much more diverse set of behavior,
  including subdiffusion, a type of anomalous diffusion exhibited by solutes in 
  this study.~\cite{klages_anomalous_2008}

@@

  %BJC4: added mention of the anomalous subdiffusion model we chose 
%  Our first modeling approach is based on the anomalous diffusion literature, 
%  applying the mathematical formalism describing subdiffusive transport to observed
%  solute trajectories. Specifically, 
  %BJC7: with changes, above thesis statement is redundant
  We treated our system as an FBM process subordinate to a CTRW, or subordinated 
  FBM (sFBM) for short. We fit sFBM
  models in two ways: First, we parameterized solute motion using a single set of parameters fit
  to a hypothesized anomalous diffusion model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's radial
  distance from the closest pore center. This allows us to include different dynamical
  behavior of solutes observed within the pore region versus within the monomer tails.

  Where applicable, the sFBM model provides a way to propagate information about
  solute trajectories gathered at the microsecond time scale to realistic experimental
  time scales. It can also confirm the importance of assumed solute-membrane interactions
  used to define the model's functional form, and suggest missing components. It
  can help researchers formulate hypotheses to explain why a certain set of parameters
  is characteristic to a specific solute.

%   is an elegant way to characterize single 
%  particle trajectories and in addition to providing an easy and efficient way to 
%  simulate solute trajectories, 
%  uncover the underlying solute-membrane interactions 
%  which result in observed transport mechanisms.

  % BJC7: moved last sentence of previous paragraph to here to make it more clear that
  % we are moving to a new model
  If one has knowledge of the primary mechanisms leading to anomalous diffusion behavior,
  one may gain additional insight by formulating a state-based model, such as Markov state
  models (MSMs). MSMs are a popular class of models used to project long timescale
  system properties based on molecular simulation trajectories by identifying
  different dynamical modes and quantifying the rates of transitions between them.
  MSMs are frequently used to study systems with slow dynamics, such as protein 
  folding.~\cite{snow_how_2005,chodera_automatic_2007} Researchers typically aim to 
  come up with a low dimensional representation of the system based on features 
  which preserve the process kinetics. This still often results in hundreds to thousands
  of distinct states.~\cite{chodera_markov_2014}

  Our second modeling approach applies an extended MSM framework to a relatively 
  small set of known states based on the three previously observed solute trapping mechanisms.  
  Standard MSMs are typically applied to determine equilibrium populations
  of states and the kinetics of transitions between those states.~\cite{bowman_using_2009} 
  We extend the framework to include state-dependent fluctuations and correlations in 
  solute position. The magnitude of solute fluctuations from its average trapped 
  position and the degree of correlation with previous fluctuations is 
  determined by the current state. To distinguish our approach from standard MSMs, we have
  named it the Markov state-dependent dynamical model (MSDDM).

  %BJC4: making goals more explicit
  We determine the degree of success of our modeling approaches in two ways. First,
  if we can closely reproduce solute MSDs measured from MD simulations with realizations
  of our models, then it is likely that the model sufficiently captures solute 
  dynamics and can be used in a predictive capacity. Even if a model fails to 
  reproduce the MD MSDs, there is value in uncovering the cause of the deviation.
  The second measure of success is by analysis of the parameters defining each 
  model. The parameters themselves should provide clues that help explain solute 
  dynamics in a coherent way. A model that successfully reproduces MD MSDs, but 
  with nonsensical parameters has severely diminished value as it could be a 
  consequence of overfitting.

@@

  We evaluate the two modeling approaches by using them to characterize the dynamical
  behavior of the four fastest moving solutes studied in our previous work:
  methanol, urea, ethylene glycol and acetic acid. In addition
  to moving quickly, allowing them to extensively explore membrane structural space,
  these solutes have a range of chemical functionality and experience each of the 
  three trapping mechanisms to different extents.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an atomistic
  molecular model of four pores in a monoclinic unit cell with 10\% water by weight 
  (see Figure~\ref{fig:membrane_structure}). Approximately one third of the water 
  molecules occupy the tail region with the rest near the pore center. We chose to
  focus our study on the 10 wt\% water system because solutes move significantly 
  faster than in the 5 wt\% system studied previously thus allowing more statistically
  valid analyses.
%  . Appropriate stochastic 
%  modeling requires that solutes sample the accessible mechanisms with representative
%  probability.

@@

  We chose to study a subset of the 4 fastest moving solutes from our previous
  work: methanol, acetic acid, urea and ethylene glycol. For each solute we 
  created a separate system and to each system we added 6 solutes per pore 
  for a total of 24 solutes. On the time scales which we simulate, this number
  of solutes per pore provides a a sufficient amount of data from which to 
  generate statistics. It also maintains a low degree of interaction between
  solutes since, at present, we are primarily interested in solute-membrane 
  interactions.
%  balance
%  of a low degree of interaction between solutes, since we are primarily interested 
%  in solute-membrane interactions at present, as well as a sufficient amount of 
%  data from which to generate statistics on the time scales which we simulate.
  Further details on the setup and equilibration of these systems are described in
  our previous work.\cite{coscia_chemically_2019}

  We extended the 1 $\mu$s simulations of our previous work to 5 $\mu$s in order
  to collect ample data. We ran MD simulations using the leapfrog integrator with 
  h-bonds constrained by the LINCS algorithm. We simulated a time step of 2 fs at
  a pressure of 1 bar and 300 K controlled by the Parrinello-Rahman barostat and
  the v-rescale thermostat respectively. We recorded frames every 0.5 ns.

@@

  %MRS6: next paragraph could be clearer.  Maybe motivate at the beginning why the importance of time. vs. ensemble averated. It's sprung a sort of a surprise there's a difference 
  %BJC6: reorganized a little and broke into paragraphs meant to convey separate ideas.
  %MRS7: definitely improved, I had a couple of comments to further refine.
  One can measure MSD in two ways. The time-averaged MSD measures all observed
  displacements over time lag $\tau$: 
  \begin{equation}
  \overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. The ensemble averaged MSD measures
  displacements with respect to a particle's initial position:
  \begin{equation}
  \langle z^2(t) \rangle = \langle z(t) - z(0) \rangle^2
  \label{eqn:ensemble_msd}
  \end{equation}
  Fits to the \textit{ensemble} averaged MSD will always reproduce the form of 
  Equation~\ref{eqn:msd_form}.

@@

  \textit{Dwell time distributions:} For subdiffusive transport, the distribution 
  of dwell times is expected to fit a power law distribution 
  proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  Because we are limited to taking measurements at discrete intervals dictated by the output 
  frequency of our simulation trajectories, we fit the empirical dwell times
  to a discrete power law distribution whose maximum likelihood $\alpha$ 
  parameter we calculated by maximizing the likelihood function: 
  %BJC6: I could move these MLE equations to Supporting  
  %MRS7: for PRE, can leave in.
  \begin{equation}
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, t_{min}) -
	\beta\sum_{i=1}^{n} \ln t_i 
  \label{eqn:powerlaw_likelihood}
  \end{equation}
  where $\beta = 1 + \alpha$, $t_i$ are collected dwell time data points,
  $n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
  where $t_{min}$ is the smallest measured value of	$t$.~\cite{clauset_power-law_2009}

@@

  \textit{Correlated hop length distributions:} The marginal distribution of hop 
  lengths by solutes undergoing an sFBM process is Gaussian, therefore we 
  parameterize it by its standard deviation, $\sigma$.
  ~\cite{metzler_random_2000, metzler_anomalous_2014,neusius_subdiffusion_2009}
  The measured mean of the hop length distribution is always very close to zero so
  we assume that it is exactly zero in our time series simulations since we have 
  no reason to expect drift in either direction based on pore symmetry.

  % BJC3: Making a logical argument regarding how we estimate H here is important
  % because I haven't seen anyone do it this way.
  sFBM implies that hops are correlated which we describe using the Hurst parameter, $H$. 
  The autocovariance function of hop lengths has the analytical form:~\cite{mandelbrot_fractional_1968}
  \begin{equation}
    \gamma(k) = \dfrac{\sigma^2}{2}\bigg[|k-1|^{2H} - 2|k|^{2H} + |k+1|^{2H}\bigg]
  \label{eqn:fbm_autocorrelation}
  \end{equation}
  where $\sigma^2$ is the variance of the underlying Gaussian distribution from which hops are
  drawn and $k$ is the time lag, or number of increments between hops. The hop autocorrelation
  function is simply Equation~\ref{eqn:fbm_autocorrelation} normalized by the variance. 
  When $H < 0.5$, hops are negatively correlated, when $H = 0.5$ we recover Brownian motion
  and when $H > 0.5$, one observes positive correlation between hops. 
  %BJC6: I could move the following to supporting and keep last sentence.
  %MRS7: could simplify down to saying that it's hard to estimate when H > 0.5, but you only have to deal with H < 0.5, so the difficulties are not relevant. 
  There are many methods
  for estimating the Hurst parameter for a time series.~\cite{clegg_practical_2006} It can
  be a difficult task because Equation~\ref{eqn:fbm_autocorrelation} decays slowly to zero, 
  especially when $H > 0.5$, meaning one needs to study large time lags with high frequency.
  Fortunately (from a mathematical perspective), all of our solutes show anti-correlated motion, so most of the information in
  Equation~\ref{eqn:fbm_autocorrelation} is contained within the first few time lags (see 
  Figure~\ref{S-fig:hurst_autocorrelation} of the Supporting Information). Any hope of
  fitting longer time lags to our data is lost in the noise since the analytical values
  are so close to zero. Therefore, we obtained H by performing a least squares fit of 
  Equation~\ref{eqn:fbm_autocorrelation} to the first 
  ten  
  0.5 ns time lags of the 
  empirically measured autocorrelation function.

@@

  \section{Results and Discussion}
